---
title: "HomeWork 3"
author: "William Mechler"
date: "3/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, loadPackages, message=FALSE, echo=FALSE}
library(knitr)
library(ggfortify)
library(knitr)
library(ggplot2)
library(gridExtra)
library(kableExtra)
library(plotly)
library(emmeans)
library(GGally)
library(tidyverse)
library(lindia)
```


```{r}
riceData <- read.table("rice.txt", header = TRUE) %>%
  select(moistureContent, amyloseContent, proteinContent, waterAbsorption, swellingRatio, pH, solidsContent, chewiness)
```
### Part 1,
Reading in the data for this problem but only selecting needed columns shown in code block above.
<br/>
<br/>
```{r, echo=FALSE}
ggpairs(riceData) +
  theme_bw()
```
<br/>
<h3> Part 2, </h3> 
From the above scatterplot matrix there appears to be some trends we can see. One of these trends is that most of our data points seems to have a low moisture content and a high amylose content. Besides that any other patterens I can see seem to be too light to call out at this point.
<br/>
<h3> Part 3, </h3> 
Below you can see the code that fits a multiple regression modeling the chewiness as a function of the following variables moisture content, amylose content, protein content, water absorption, swelling ratio, pH, and solids content.
<br/>
```{r, warning=FALSE}
riceFit_part3 <- lm(chewiness ~ moistureContent + amyloseContent + proteinContent + waterAbsorption + swellingRatio + pH + solidsContent, data=riceData)

autoplot(riceFit_part3)
```

<h3> Part 4, </h3>
It appears our model right now violates normally and constant variance. We can see these violates within out Normal Q-Q plot and how in our Residuals vs Fitted it was a scale from -250 to around 250 which again would violate the constant variance assumption.
<br/>
```{r, warning=FALSE}
gg_boxcox(riceFit_part3)
```
<h3>Part 5,</h3> 
The above box cox shows a optimal λ of -0.5, this would be a lambda of -0.5 would be applied as a negative square root. Also with -1.7 and 0.6 being at the edge of our box cox we could also try a normal square root transformation and with the -1.7 put chewiness to the power of -1.7. Also with λ = 0 withing our box cox we could also do a transformation of log of chewiness.
<br/>


```{r, warning=FALSE}
riceFit_Part6 <- lm(log10(chewiness) ~ moistureContent + amyloseContent + proteinContent + waterAbsorption + swellingRatio + pH + solidsContent, data=riceData)

autoplot(riceFit_Part6)
```
<br/>
<h3> Part 6, </h3>
above you can see the code that fits a multiple regression modeling the chewiness as a function of the following variables moisture content, amylose content, protein content, water absorption, swelling ratio, pH, and solids content. However we applied a log transformation as λ = 0 is within our confidence band.

<h3> Part 7, </h3>
Our log10 transformation fixed our Normal Q-Q plot which helped with our normality assumption. Also the scale on our Residuals vs Fitted went from having a scale from around -250 to 250 to after our transformation of having a scale of -0.2 to .02. This specifialy helped fixed our problem with the constant variance assumption.
<br/>

```{r, warning=FALSE}
summary(riceFit_Part6)
```
<h3> Part 8, </h3> 
From our model we can say when variables are zero our mean chewiness would be -2.9353164 which wouldn't make sense because you can't have a negative chewiness level.

<h3> Part 9, </h3>
The slope coefficients for moisture content and protein content are 0.1989953 and 0.1559255 respectably. This means for moisture content if all other predictor variables are held constant a 1 unit increase in moisture content would increase chewiness by 0.1989953. Also for protein content if we held all other predictor variables constant a 1 unit increase in protein content would increase chewiness by 0.1559255.

<h3> Part 10, </h3>
Our model from part 6 is significant with a p-value of 0.007886 and with an F-statistic of 3.763.
<br/>

```{r, warning=FALSE}
fictional_rice <- data.frame(moistureContent = 14.9, 
                               amyloseContent = 22.6,
                               proteinContent = 8.1, 
                               waterAbsorption = 211,
                               swellingRatio = 225,
                               pH = 6.6,
                               solidsContent   = 5.2)

predict(riceFit_Part6, newdata=fictional_rice, interval="prediction", level=0.95)
```
<h3> Part 11, </h3>
From the above code we see that our new model from part 6 would predict our fictional rice to have a chewiness level from 2.56 to 3.10 with a 95% confidence level. 
<br/>
