---
title: "HomeWork 3"
author: "William Mechler"
date: "3/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, loadPackages, message=FALSE, echo=FALSE}
library(knitr)
library(ggfortify)
library(knitr)
library(ggplot2)
library(gridExtra)
library(kableExtra)
library(plotly)
library(emmeans)
library(GGally)
library(tidyverse)
library(lindia)
```

<br/>
Part 1, reading in the data for this problem but only selecting needed columns.
<br/>
```{r}
riceData <- read.table("rice.txt", header = TRUE) %>%
  select(moistureContent, amyloseContent, proteinContent, waterAbsorption, swellingRatio, pH, solidsContent, chewiness)
```

```{r, echo=FALSE}
ggpairs(riceData) +
  theme_bw()
```
<br/>
Part 2, from the above scatterplot matrix there appears to be some trends we can see. One of these trends is that most of our data points seems to have a low moisture content and a high amylose content. Besides that any other patterens I can see seem to be too light to call out at this point.
<br/>
```{r, warning=FALSE}
riceFit1 <- lm(chewiness ~ moistureContent + amyloseContent + proteinContent + waterAbsorption + swellingRatio + pH + solidsContent, data=riceData)

autoplot(riceFit1)
```
<br/>
Part 3, above you can see the code that fits a multiple regression modeling the chewiness as a function of the following variables moisture content, amylose content, protein content, water absorption, swelling ratio, pH, and solids content.

Part 4, residual analysis TBD
<br/>
```{r, warning=FALSE}
gg_boxcox(riceFit1)
```
<br/>
Part 5, The above box cox shows a optimal λ of -0.5, this would be a lambda of -0.5 would be applied as a negative square root.
<br/>
```{r, warning=FALSE, echo=FALSE}
riceFit2 <- lm(-sqrt(chewiness) ~ moistureContent + amyloseContent + proteinContent + waterAbsorption + swellingRatio + pH + solidsContent, data=riceData)

autoplot(riceFit2)
```


```{r, warning=FALSE}
riceFit_Part6 <- lm(log(chewiness) ~ moistureContent + amyloseContent + proteinContent + waterAbsorption + swellingRatio + pH + solidsContent, data=riceData)

autoplot(riceFit_Part6)
```
<br/>
Part 6, above you can see the code that fits a multiple regression modeling the chewiness as a function of the following variables moisture content, amylose content, protein content, water absorption, swelling ratio, pH, and solids content. However we applied a log transformation as λ = 0 is within our confidence band.

Part 7, residual analysis TBD
<br/>

```{r, warning=FALSE}
summary(riceFit_Part6)
```

<br/>
Part 8,

Part 9,

Part 10,
<br/>

```{r, warning=FALSE}
fictional_rice <- data.frame(moistureContent = 14.9, 
                               amyloseContent = 22.6,
                               proteinContent = 8.1, 
                               waterAbsorption = 211,
                               swellingRatio = 225,
                               pH = 6.6,
                               solidsContent   = 5.2)

predict(riceFit_Part6, newdata=fictional_rice, interval="prediction", level=0.95)
```
<br/>
Part 11,
<br/>